Visualizations for the manuscript

### 2022-05-14 metanalysis_visualization/manuscript_Figures.Rmd
reformating bootstrap sex

```{r libraries, message=FALSE, warning=FALSE, results='hide'}
# Install packages
paket <- function(pak){
  new_pak <- pak[!(pak %in% rownames(installed.packages()))]
  if (length(new_pak)) 
    install.packages(new_pak, dependencies = TRUE,repos="https://cloud.r-project.org/")
  sapply(pak, library, character.only = TRUE)
}
listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR", "scales", "EHRtemporalVariability", "plotly", "gmodels")
paket(listOfPackages)

@@

#### Trend differences

plot_annotations_df_primary <- pvalues_boot %>%
  select(siteid, label) %>% 
  filter(term == "difference_slope") %>% 
  left_join(country_map)
```

@@

```{r, include=FALSE}
patients_hospitalized_agg_sex_total <- patients_hospitalized_agg_sex_perc_all %>%
  select( - percentage ) %>%
  group_by( time_p, sex ) %>%
  summarise( total = sum( count ), 
             count_sex = sum( count_sex ), 
             percentage = 100* count_sex / total, 
             siteid = "ALL")
```

@@

```{r}
patients_hospitalized_agg_sex_total <- patients_hospitalized_agg_sex_total %>% 
  mutate(period = ifelse(time_p <= as.Date("2020-03-15"), "before_pandemic", "during_pandemic")) %>% 
  select(time_p, period, sex, count_sex, percentage, siteid)
```

@@

bootstrapped_coefficients_sex <- lapply(bootstrapped_sex, `[[`, 1) %>%
  bind_rows() %>% 
  rename(before_slope = time, 
         difference_slope = `periodduring_pandemic:time`)

@@

coefficients_df_sex <- patients_hospitalized_agg_sex_total %>% 
  group_by(sex) %>% 
  mutate(month_year = format(time_p, "%m"),
         time = seq_along(time_p))  %>% 
  group_modify(coefficient_slope_calculation, 
               slope_name = "periodduring_pandemic:time") %>% 
  rename(observed_slope_coefficient = estimate)

df_coefficients_boot_sex <- bootstrapped_coefficients_sex %>% 
  left_join(coefficients_df_sex, by = c("sex"))
```

### Alternative plot

```{r }
names_sex <- unique(df_coefficients_boot_sex$sex)

plot_annotations_sex <- df_coefficients_boot_sex %>% 
  group_by(sex) %>% 
  group_map( bootstrap_pvalue_calculation, slope_name = "difference_slope" ) %>%
  set_names(names_sex) %>% 
  lapply(FUN = function(vec) {
    paste0("Coeffs differences: ",
           vec$observed_slope_coefficient,
           " (pvalue: ",
           format.pval(vec$pvalue, 
                       eps = 10**-3),
           ")")
  }) %>% 
  unlist()
plot_annotations_sex_df <- data.frame(
  sex = names(plot_annotations_sex), 
  label = plot_annotations_sex
)
```

```{r}
bootstrapped_fitted_sex %>% 
  mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset")) %>% 
  ggplot(aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
  facet_wrap("sex",
             scales = "free_y",
             labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
  geom_line(aes(color = period),
            alpha = 0.03) +
  geom_line(data = quantile_bootstrapped_fitted_sex,
            aes(y = value, group = interaction(quantiles, period)),
            linetype = "dashed") +
  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
            aes(y = value, group = period)) +
  geom_point(data = patients_hospitalized_agg_sex_total,
             aes(y = percentage, group=NULL),
             alpha = 0.5) +
  scale_colour_manual(values = cbPalette) +
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
  guides(color = guide_legend(override.aes = list(alpha = 1))) +
  geom_label(data = plot_annotations_sex_df, 
             aes(label=label), 
            x = -Inf,
            y = +Inf,
            hjust = 0,
            vjust = 1,
            size = 4,
            inherit.aes = FALSE) +
  labs(x = "Calendar dates",
       y = "Percentage",
       color = "Period") +
  theme(strip.text = element_text(size = 10),
        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
        axis.text.y = element_text(size = 8), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 14),
        plot.subtitle = element_text(size = 8, face = "italic"), 
        legend.position = "bottom") 
ggsave("figure4_boostrapped_sex.png", height = 15, width = 25, units = "cm")
```

@@

```{r annotation bootstrap disorder sex}
names_sex <- unique(df_coefficients_boot_sex$sex)

## Using only male, but female and male necessarily yield equal p-values

pvalue_sex_slope_diff <- bootstrap_pvalue_calculation(
  df_coefficients_boot_sex[df_coefficients_boot_sex$sex == "female", ], 
  slope_name = "difference_slope"
)

plot_annotations_sex_df <- paste0("Female proportion period differences: ",
                                  pvalue_sex_slope_diff$difference_slope,
                                  " (pvalue: ",
                                  format.pval(pvalue_sex_slope_diff$pvalue,
                                              digits = 2,
                                              eps = 10**-3, 
                                              nsmall = 2),")") %>%
  data.frame(label = ., 
             plotsLabel = "Female")
```

```{r bootstrap_plot}
ggplot(bootstrapped_fitted_sex,
       aes(x = time_p, y = predicted/100, group = interaction(iteration, period, sex))) +
  geom_line(aes(color = interaction(period, sex)),
            alpha = 0.01) +
  geom_line(data = quantile_bootstrapped_fitted_sex,
            aes(y = value/100, group = interaction(quantiles, period, sex)),
            linetype = "dashed") +
  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
            aes(y = value/100, group = interaction(period, sex))) +
  geom_point(data = patients_hospitalized_agg_sex_total,
             aes(y = percentage/100, group=sex),
             alpha = 1) +
  # scale_color_brewer(palette = "Paired") +
  scale_color_manual(labels = c("Female", "Male", "", ""),
                     values = c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C")) +
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
  guides(colour = guide_legend(
    override.aes = list(alpha = 1,
                        colour = c('#1F78B4', '#33A02C', NA, NA))) 
  ) +
  geom_label(data = plot_annotations_sex_df,
             aes(label = label),
             x = -Inf,
             y = +Inf,
             hjust = 0,
             vjust = 1,
             size = 4,
             inherit.aes = FALSE) +
  labs(title = "Interrupted Time Series Analysis: Patients sex proportion",
       subtitle = "Parametric bootstrap estimation: (10000 iterations)",
       x = "Calendar dates",
       y = "Sex proportion",
       color = "Sex") +
  theme(strip.text = element_text(size = 10),
        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
        axis.text.y = element_text(size = 8),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 14),
  plot.subtitle = element_text(size = 12, face = "italic"))

ggsave("figure4_ITS_sex.png", height = 20, width = 20, units = "cm")
```

>>>>

```{r libraries, message=FALSE, warning=FALSE, results='hide'}
# Install packages
paket <- function(pak){
  new_pak <- pak[!(pak %in% rownames(installed.packages()))]
  if (length(new_pak)) 
    install.packages(new_pak, dependencies = TRUE,repos="https://cloud.r-project.org/")
  sapply(pak, library, character.only = TRUE)
}
listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR", "scales", "EHRtemporalVariability", "plotly", "gmodels", "flextable")
paket(listOfPackages)

@@

#### Primary analysis plot

```{r}
plot_annotations_df_primary <- pvalues_boot %>%
  select(siteid, label) %>% 
  filter(term == "difference_slope") %>% 
  left_join(country_map)
```

@@

```{r, include=FALSE}
patients_hospitalized_agg_sex_total <- patients_hospitalized_agg_sex_perc_all %>%
  select( - percentage ) %>%
  group_by( time_p, sex ) %>%
  summarise( total = sum( count ), 
             count_sex = sum( count_sex ), 
             percentage = 100* count_sex / total, 
             siteid = "ALL") %>%  
  mutate(period = ifelse(time_p <= as.Date("2020-03-15"), "before_pandemic", "during_pandemic")) %>% 
  select(time_p, period, sex, count_sex, percentage, siteid)
```

```{r}
observed_coefficients_df_sex <- patients_hospitalized_agg_sex_total %>% 
  group_by(sex) %>% 
  mutate(month_year = format(time_p, "%m"),
         time = seq_along(time_p))  %>% 
  group_modify(coefficient_slope_calculation, 
               slope_name = c("time", "periodduring_pandemic:time")
               ) %>% 
  rename(observed_slope_coefficient = estimate) %>% 
  pivot_wider(names_from = term,
              values_from = observed_slope_coefficient) %>% 
  rename(before_slope = time, 
         difference_slope = `periodduring_pandemic:time`) %>% 
  mutate(during_slope = before_slope + difference_slope) %>% 
  pivot_longer(cols = c(before_slope, difference_slope, during_slope), 
               names_to = "term", 
               values_to = "observed_slope_coefficient")

@@

@@

bootstrapped_coefficients_sex <- lapply(bootstrapped_sex, `[[`, 1) %>%
  bind_rows() %>% 
  rename(before_slope = time, 
         difference_slope = `periodduring_pandemic:time`) %>% 
  mutate(during_slope = before_slope + difference_slope) %>% 
  select(iteration, sex, everything()) %>% 
  pivot_longer(cols = c(before_slope, difference_slope, during_slope), 
               names_to = "term", 
               values_to = "bootstrapped_slope_coefficient") %>% 
  left_join(observed_coefficients_df_sex, by = c("sex", "term"))

df_coefficients_boot_sex <- bootstrapped_coefficients_sex %>% 
  left_join(coefficients_df_sex, by = c("sex"))

pvalues_boot_sex <- bootstrapped_coefficients_sex %>% 
  group_by(sex, term) %>% 
  group_modify( bootstrap_pvalue_calculation, .keep = TRUE) %>% 
  mutate(label = paste0("Female proportion period differences: ",
                        observed_slope_coefficient,
                        " (pvalue: ",
                        format.pval(pvalue, digits = 2, eps = 10**-2), ")")
  )

ci_boot_sex <- bootstrapped_coefficients_sex %>% 
  group_by(sex, term) %>% 
  summarise("2.5%" = quantile(bootstrapped_slope_coefficient, 0.025), 
            "50%" = quantile(bootstrapped_slope_coefficient, 0.50), 
            "97.5%" = quantile(bootstrapped_slope_coefficient, 0.975)) %>% 
  mutate(ci = paste0("[", as.character(round(`2.5%`, 2)), "; ", as.character(round(`97.5%`, 2)), "]"))

@@

#### Table with bootstrapped coefficients, and pvalues

```{r}
pvalues_boot_sex %>% 
  inner_join(ci_boot_sex) %>% 
  arrange(sex) %>% 
  mutate(string = paste0(observed_slope_coefficient, " ", ci, " (", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
  select(sex, string) %>% 
  pivot_wider(names_from = term, values_from = string) %>% 
  select(sex, before_slope, during_slope, difference_slope) %>% 
  rename("Before pandemic trend [95% CI] (p-value)" = before_slope,
        "Since pandemic trend [95% CI] (p-value)" = during_slope,
        "Difference before/since pandemic trend [95% CI] (p-value)" = difference_slope) %>% 
  flextable::flextable() %>% 
  flextable::save_as_docx(path = "table_sex.docx")

@@

```{r}
plot_annotations_sex_df <- pvalues_boot_sex %>%
  select(term, sex, label) %>% 
  filter(term == "difference_slope" & sex == "female") %>% 
  mutate(plotsLabel = "Female")

```{r bootstrap_plot}
ggplot(bootstrapped_fitted_sex,
       aes(x = time_p, y = predicted/100, group = interaction(iteration, period, sex))) +
  geom_line(aes(color = interaction(period, sex)),
            alpha = 0.01) +
  geom_line(data = quantile_bootstrapped_fitted_sex,
            aes(y = value/100, group = interaction(quantiles, period, sex)),
            linetype = "dashed") +
  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
            aes(y = value/100, group = interaction(period, sex))) +
  geom_point(data = patients_hospitalized_agg_sex_total,
             aes(y = percentage/100, group=sex),
             alpha = 1) +
  # scale_color_brewer(palette = "Paired") +
  scale_color_manual(labels = c("Female", "Male", "", ""),
                     values = c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C")) +
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
  guides(colour = guide_legend(
    override.aes = list(alpha = 1,
                        colour = c('#1F78B4', '#33A02C', NA, NA))) 
  ) +
  geom_label(data = plot_annotations_sex_df,
             aes(label = label),
             x = -Inf,
             y = +Inf,
             hjust = 0,
             vjust = 1,
             size = 4,
             inherit.aes = FALSE) +
  annotate("text_female", x = 7, y = 0.85, label = "Female proportion") + 
  annotate("text_male", x = 7, y = 0.15, label = "Male proportion") + 
  labs(title = "Interrupted Time Series Analysis: Patients sex proportion",
       subtitle = "Parametric bootstrap estimation: (10000 iterations)",
       x = "Calendar dates",
       y = "Sex proportion",
       color = "Sex") +
  theme(strip.text = element_text(size = 10),
        axis.text.x=element_text(angle = 45, hjust=1, size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 14),
  plot.subtitle = element_text(size = 12, face = "italic"))

ggsave("figure4_ITS_sex.png", height = 15, width = 20, units = "cm")
```

#### Alternative plot 
```{r, include=FALSE}
bootstrapped_fitted_sex %>% 
  mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset")) %>% 
  ggplot(aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
  facet_wrap("sex",
             scales = "free_y",
             labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
  geom_line(aes(color = period),
            alpha = 0.03) +
  geom_line(data = quantile_bootstrapped_fitted_sex,
            aes(y = value, group = interaction(quantiles, period)),
            linetype = "dashed") +
  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
            aes(y = value, group = period)) +
  geom_point(data = patients_hospitalized_agg_sex_total,
             aes(y = percentage, group=NULL),
             alpha = 0.5) +
  scale_colour_manual(values = cbPalette) +
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
  guides(color = guide_legend(override.aes = list(alpha = 1))) +
  geom_label(data = plot_annotations_sex_df, 
             aes(label=label), 
            x = -Inf,
            y = +Inf,
            hjust = 0,
            vjust = 1,
            size = 4,
            inherit.aes = FALSE) +
  labs(x = "Calendar dates",
       y = "Percentage",
       color = "Period") +
  theme(strip.text = element_text(size = 10),
        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
        axis.text.y = element_text(size = 8), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 14),
        plot.subtitle = element_text(size = 8, face = "italic"), 
        legend.position = "bottom") 
ggsave("figure4_boostrapped_sex.png", height = 15, width = 25, units = "cm")
```
